---
title: "Class 1 methods 3"
author: "Caroline Hommel"
date: "9/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 3) Brushing up on the General Linear Model

We'll do a light start and get you back in the game of thinking about formulae and how to build your linear models  
Finally, we'll have a few exercises, finishing off today's practical exercises 

## A list of formulae
```{r, eval=FALSE}
formula <- y ~ x ## y as a function of x
y ~ 1 ## model the intercept for "y"
y ~ x ## model the main effect of x and the intercept for y
y ~ x + 1 ## the same as above (+ 1 is implicit)
y ~ x + 0 ## model the main effect of x and no intercept
y ~ x - 1 ## the same as above
y ~ 0 ## doesn't model anything (for completeness)
y ~ x + z ## model the main effects x and z (and an intercept)
y ~ x:z ## model interaction of x and z
y ~ x * z ## model the main effects x and z and their interaction
y ~ x + z + x:z ## the same as above
```

## Dataset mtcars
Let's look at the "mtcars" data:  

_[, 1]   mpg   Miles/(US) gallon  
[, 2]	 cyl	 Number of cylinders  
[, 3]	 disp	 Displacement (cu.in.)  
[, 4]	 hp	 Gross horsepower  
[, 5]	 drat	 Rear axle ratio  
[, 6]	 wt	 Weight (lb/1000)  
[, 7]	 qsec	 1/4 mile time  
[, 8]	 vs	 V/S  
[, 9]	 am	 Transmission (0 = automatic, 1 = manual)  
[,10]	 gear	 Number of forward gears  
[,11]	 carb	 Number of carburetors_  


## Miles per gallon and weight

We can do a scatter plot, and it looks like there is some relation between fuel usage and the weight of cars.
Let's investigate this further

```{r,fig.height=5, fig.width=6}
par(font.lab=2, font.axis=2, cex=1.2)
plot(mpg ~ wt, data=mtcars, xlab='Weight (lb/1000)', ylab='Miles/(US) gallon',
     main='Scatter plot', ylim=c(0, 40))
```

# Exercises and objectives
The objectives of today's exercises are:  
1) To remind you of the (general) linear model, and how we can use it to make models in R  
2) To make some informal model comparisons  
3) To estimate models based on binomially distributed data  

If you would like to read more about a given function, just prepend the function with a question mark, e.g.  
``` {r, eval=FALSE}
?lm
```

REMEMBER: In your report, make sure to include code that can reproduce the answers requested in the exercises below   

## Exercise 1
The general linear model: $Y = X \beta + \epsilon$:  
Do a linear regression, expressing fuel usage as a function of weight using the function __lm__  
```{r, eval=FALSE}
data(mtcars)
model <- lm( mpg ~ wt, data=mtcars)
summary(model)
```
1. extract $\hat{\beta}$, $Y$, $\hat{Y}$, $X$ and $\epsilon$ from __model__ (hint: have a look at the function __model.matrix__)  
    
```{r}
model.matrix(model)

#estimates intersept and slope
beta_hat <- model$coefficients

#hat is an estimate. y = intercept + slope * x (y=ax+b)
Y_hat <- beta_hat[1] + mtcars$wt*beta_hat[2]

# my x and y's are the miles pr gallon and weight for my linear model
Y <- mtcars$mpg
X <- mtcars$wt

# we compare what it has estimated and what the real datapoints are 
error <- model$residuals
```
     i. create a plot that illustrates $Y$ and $\hat{Y}$ (if you are feeling ambitious, also include $\epsilon$ (hint: you can use the function __arrows__))
```{r}
plot(X, Y)
abline(a = beta_hat[1], b = beta_hat[2], col = "orange")
arrows(x0=X, y0= Y, y1 = Y_hat, length = 0, code = 2, angle = 0)
```
The green is the predicted linear line and the dots are the actual/observed values of Y
     
2. estimate $\beta$ for a quadratic model ($y = {\beta}_{2} x^2 + {\beta}_{1} x + {\beta}_{0}$) using ordinary least squares _without_ using __lm__; $\hat{\beta} = {({X}^{T} X)}^{-1} {X}^{T} Y$ (hint: add a third column to $X$ from step 1)
```{r}
n <- 32
const <- rep(1,n)
design_matrix <- cbind(const, mtcars$wt, (mtcars$wt)^2)
#udregner intercept og to forskellige slopes. 
beta_ols <- solve(t(design_matrix)%*%design_matrix)%*%t(design_matrix)%*%Y

```

3. compare your acquired $\hat{\beta}$ with the output of the corresponding quadratic model created using __lm__ (hint: use the function __I__, see details under help and the sub-section formula operators here: https://www.datacamp.com/community/tutorials/r-formula-tutorial)  

```{r}
#creating a new lm with the Y predicted by the design matrix (which has the weight and the weight^2 in two columns)
lm_compare <- lm(Y ~ design_matrix [,2] + design_matrix [,3])

summary(lm_compare)

```
The values from the ols_beta (49.93 and -13.38 and 1.17) are the same as those from the second lm calculated with the design matrix. So the lm function uses the formula I used in the beta_ols and that is why the values are the same. One thing we calculated 'manually' the other with lm.


i. create a plot that illustrates $Y$ and $\hat{Y}$ (if you are feeling ambitious, also include $\epsilon$ (hint: you can use the function __arrows__)) 
    
```{r}
#now we will compare some predicted y-values with the observed ones from the mtcars data set
design_matrix <- as.matrix(design_matrix)
yhat <- design_matrix%*%beta_ols

plot(Y, yhat)
abline(a = 0, b = 1, col = "red")
# a=intercept and b er hældning, så laver bare en hel standard linje for at sammenligne predictions med the observed values. Havde de været perfekte havde de self fulgt den orange flotte lige linje. 
```

## Exercise 2
Compare the plotted quadratic fit to the linear fit  
```{r}
plot(Y, Y_hat)
abline(a=0, b=1, col="green")

plot(Y, yhat)
abline(a=0, b=1, col="yellow")
```
1. Which seems better?  
The quadratic (number 2) seems to catch the tendencies of the data a little bit better than the first one. This can be seen as the errors are more evenly distributed along the line in the quadratic plot. 


2. calculate the sum of squared errors, (show the calculation based on $\epsilon$). Which fit has the lower sum?  
```{r}
sum((Y-Y_hat)^2)

sum((Y-yhat)^2)

#sum((model$residuals)^2)
```
The sum of the squared errors is lower for the quadratic model, which means that it's differences between observed values and predicted values are lower than the ones from the linear fit. Ergo the quadratic fit is better. 

3. now make a cubic fit ($y = {\beta}_{3} x^3 + {\beta}_{2} x^2 + {\beta}_{1} x + {\beta}_{0}$) and compare it to the quadratic fit  
```{r}
cubic_fit <- cbind( design_matrix, (mtcars$wt)^3)

#added cubic fit to the model in a new column
lm_cubic_compare <- lm(Y ~ design_matrix [,2] + design_matrix [,3] + cubic_fit [,4])

```   

   i. create a plot that illustrates $Y$ and $\hat{Y}$ for both the cubic and the quadratic fits (plot them in the same plot)
```{r}
both_plot <- ggplot(mtcars, aes(x = wt, y = mpg))+
  geom_point()+
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 1, se = FALSE)+
  stat_smooth(method = "lm", formula = y ~ x + I(x^2) + I(x^3), size = 1, color = "green", se = FALSE)
both_plot
```
    ii. compare the sum of squared errors  
```{r}
sum((Y-yhat)^2)
sum((lm_cubic_compare$residuals)^2)
```
The cubic fit is a tiny bit lower and thereby better

  iii. what's the estimated value of the "cubic" (${\beta}_3$) parameter? 
```{r}
coef(lm_cubic_compare)
```
  It's a small number, so it doesn't have great influence on the model. (0.04593)


4. bonus question: which summary statistic is the fitted value (_Intercept_ or ${\beta}_0$ in $y = {\beta}_0$) below identical to?
```{r, echo=FALSE}
lm(mpg ~ 1, data=mtcars)
```

## Exercise 3
Doing a logistic regression - estimating the probability that a car has automatic transmission (0) or manual transmission (1) based on its weight
```{r, eval=FALSE}
data(mtcars)
logistic.model <- glm(formula= am ~ wt, data=mtcars, family='binomial')
```

Probabilities live on the range $(0, 1)$ - using the so-called logit function as a "link-function" we can map these onto the range $(-\infty, \infty)$, i.e. the real numbers.  

What we model in this case is: $Pr(y = 1) = logit^{-1}(X \beta)$, i.e. the probability of a car having manual transmission, given its weight. $X \beta$ is called the linear predictor; compare with $Y = X \beta + \epsilon$ 
It is helpful to define the logit function and its inverse function for the following:  

```{r}
logit <-     function(x) log(x / (1 - x))
inv.logit <- function(x) exp(x) / (1 + exp(x))
```

1. plot the fitted values for __logistic.model__:  
    i. what is the relation between the __linear.predictors__ and the __fitted_values__ of the __logistic.model__ object?
2. plot the logistic function, you've estimated based on your $\hat{\beta}$, (not just the fitted values). Use an _xlim_ of (0, 7)
    i. what's the interpretation of the estimated $\hat{\beta}_0$ (the _Intercept_)
    ii. calculate the estimated probability that the Pontiac Firebird has automatic transmission, given its weight
    iii. bonus question - plot the logistic function and highlight all the cars where we guessed wrongly, if we used the following "quantizer" function:
```{r}
getwd()
```
    
\begin{equation}
  transmission_{guess}=
  \begin{cases}
    1 (manual), & \text{if}\ PR(y = 1) ≥ 0.5 \\
    0 (automatic), & \text{otherwise}
  \end{cases}
\end{equation}    
    
3. plot quadratic fit alongside linear fit  
    i. judging visually, does adding a quadratic term make a difference?
    ii. check the details in the help of the AIC function - which of the models provide the better fit according to the AIC values and the residual deviance respectively?
    iii. in your own words, why might it be good to penalise a model like the quadratic model, we just fitted.
    
# Next time
We are going to looking at extending our models with so called random effects. We need to install the package "lme4" for this. Run the code below or install it from your package manager (Linux)  
```{r, eval=FALSE}
install.packages("lme4")
```
We can fit a model like this:

```{r}
library(lme4)
mixed.model <- lmer(mpg ~ wt + (1 | cyl), data=mtcars)
```

They result in plots like these:
```{r}
par(font.lab=2, font.axis=2, cex=1.2)
plot(mtcars$wt, fitted.values(mixed.model),
     main='Linear regression with group intercepts (n cylinders)',
    xlab='Weight (lb/1000)', ylab='Miles/(US) gallon',
    pch=3)
```

and this
```{r}
mixed.model <- lmer(mpg ~ wt + (wt | cyl), data=mtcars)
plot(mtcars$wt, fitted.values(mixed.model),
     main='Linear regression with group intercepts and group slopes (n cylinders)',
    xlab='Weight (lb/1000)', ylab='Miles/(US) gallon',
    pch=3)
``` 

but also new warnings like:  

Warning:
In checkConv(attr(opt, "derivs"), opt\$par, ctrl = control$checkConv,  :
  Model failed to converge with max|grad| = 0.0121962 (tol = 0.002, component 1)

